---
title: "Lab 07 - Simpson's paradox"
author: "John Adams"
date: "3/15/21"
output: github_document
---

### Directions for Lab
https://rstudio-education.github.io/datascience-box/course-materials/lab-instructions/lab-07/lab-07-simpsons-paradox.html

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(mosaicData) 
```

### Exercise 1


```{r the-data}
?Whickham
```

Observational study. 

### Exercise 2

```{r understand-data}
head(Whickham)
nrow(Whickham)
glimpse(Whickham)
```

1,314 people were included in this survey. 

### Exercise 3

The researchers recorded 3 variables: age at time of first survey, whether or not they were a smoker, and if the person was alive or dead 20 years after the first survey.

### Exercise 4

I'd expect poor health outcomes for those that smoke. 

### Exercise 5

```{r table}
Whickham %>%
ggplot(aes(x = smoker, fill = outcome)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c(
    "Alive" = "#2166ac",
    "Dead"  = "#b2182b"
  )) + 
  labs(title = "Relationship between smoking status and health outcomes",
       subtitle = "1,314 Women first surveyed between 1972-1974 and followed-up with 20 years later",
       x = "Whether or not the woman smoked", y = "Proportion", 
       fill = "Health outcome after 20 years") +
  theme_minimal()

Whickham %>%
  count(smoker, outcome) %>% 
  group_by(smoker) %>%
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>%
  pivot_wider(names_from = smoker, values_from = proportion)
```
This visualization shows that those who did not smoke at the time of the survey had a higher rate of death after 20 years. 
```{r age-cats}
Whickham_cat <- Whickham %>%
  mutate(
    age_cat = case_when(
      age <= 44 ~ "18-44",
      age > 44 & age <= 64 ~ "45-64",
      age > 64 ~ "65+"
    )
  )
```
```{r age-cats-viz}

Whickham_cat %>%
  ggplot(aes(x = smoker, fill = outcome)) +
  geom_bar(position = "fill") +
  facet_wrap(~age_cat, nrow = 1, ncol = 3) +
  theme_minimal() +
  scale_fill_manual(values = c(
    "Alive" = "#2166ac",
    "Dead"  = "#b2182b"
  )) + 
  labs(title = "Relationship between smoking status and health outcomes by age group",
       subtitle = "1,314 women first surveyed between 1972-1974 and followed-up with 20 years later",
       x = "Whether or not the woman smoked", y = "", 
       fill = "Health outcome after 20 years")

```
```{r}
Whickham_cat %>%
  count(smoker, age_cat, outcome) %>% 
  group_by(smoker, age_cat) %>%
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>%
  pivot_wider(names_from = age_cat, values_from = proportion)
```

In contrast to our first visualization, these bar charts show that within each age group, the rate of death is higher for those that smoked. The oldest age group, 65+, has such high rates of death after 20 years, that they are causing the overal to paint an incorrect picture, hense the simpsons paradox. 


