---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "John Adams"
date: "3/14/21"
output: github_document
---

### Lab Directions
https://rstudio-education.github.io/datascience-box/course-materials/lab-instructions/lab-05/lab-05-wrangle-sp-data.html

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
library(skimr)
```

```{r load-data, message = FALSE}
states <- read_csv(url("https://raw.githubusercontent.com/rstudio-education/datascience-box/master/course-materials/starters/lab/lab-04-viz-sp-data/data/states.csv"))
```

### Exercise 1

```{r AK-dn-data}
dn_ak <- dennys %>%
  filter(state == "AK")

nrow(dn_ak)
```


### Exercise 2

```{r AK-lq-data}
lq_ak <- laquinta %>%
  filter(state == "AK")

nrow(lq_ak)
```



### Exercise 3

```{r join-ak}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```


### Exercise 4
```{r view-ak-join}
glimpse(dn_lq_ak)
nrow(dn_lq_ak)
```

### Exercise 5
```{r haversine-function}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}


```

### Exercise 6

```{r distance}
dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```


### Exercise 7

```{r min-dist}

dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarise(closest = min(distance)) %>%
  arrange(-desc(closest))
```

### Exercise 8

```{r distribution-min-dist}
dn_lq_ak_mindist %>%
  ggplot(aes(x = closest)) +
  geom_histogram() +
  labs(title = "Shortest Distances Between Denny's and La Quinta Hotels",
       subtitle = "in Alaska",
       x = "Distance (km)") +
  theme_minimal()

dn_lq_ak_mindist %>%
  skim(closest)
```
### Exercise 9

filter each datat frame to just NC and then join. 
```{r nc-dn-lq}

dn_nc <- dennys %>%
  filter(state == "NC")

nrow(dn_nc)

lq_nc <- laquinta %>%
  filter(state == "NC")

nrow(lq_nc)

12*28 #prediction of rows

dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")
dn_lq_nc
```
```{r add-distance-nc}
dn_lq_nc <- dn_lq_nc %>%
   mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```
```{r min-dist-nc}

dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarise(closest = min(distance)) %>%
  arrange(-desc(closest))
```


```{r distribution-min-dist-nc}
nc_hist <- dn_lq_nc_mindist %>%
  ggplot(aes(x = closest)) +
  geom_histogram(binwidth = 10) +
  labs(title = "Shortest Distances Between Denny's and La Quinta Hotels",
       subtitle = "in North Carolina",
       x = "Distance (km)") +
  theme_minimal()

dn_lq_nc_mindist %>%
  skim(closest)
```

### Exercise 10

filter each data frame to just TX and then join. 
```{r tx-dn-lq}

dn_tx <- dennys %>%
  filter(state == "TX")

nrow(dn_tx)

lq_tx <- laquinta %>%
  filter(state == "TX")

nrow(lq_tx)

200*237 #prediction of rows

dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")
dn_lq_tx
```
```{r add-distance-tx}
dn_lq_tx <- dn_lq_tx %>%
   mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```

```{r min-dist-tx}

dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarise(closest = min(distance)) %>%
  arrange(-desc(closest))
```

```{r distribution-min-dist-tx}
tx_hist <- dn_lq_tx_mindist %>%
  ggplot(aes(x = closest)) +
  geom_histogram(binwidth = 10) +
  labs(title = "Shortest Distances Between Denny's and La Quinta Hotels",
       subtitle = "in Texas",
       x = "Distance (km)") +
  theme_minimal()

dn_lq_tx_mindist %>%
  skim(closest)
```

### Exercise 11a

filter each data frame to just MA and then join. 
```{r ma-dn-lq}

dn_ma <- dennys %>%
  filter(state == "MA")

nrow(dn_ma)

lq_ma <- laquinta %>%
  filter(state == "MA")

nrow(lq_ma)

8*6 #prediction of rows

dn_lq_ma <- full_join(dn_ma, lq_ma, by = "state")
dn_lq_ma
```
```{r add-distance-ma}
dn_lq_ma <- dn_lq_ma %>%
   mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```

```{r min-dist-ma}

dn_lq_ma_mindist <- dn_lq_ma %>%
  group_by(address.x) %>%
  summarise(closest = min(distance)) %>%
  arrange(-desc(closest))
```

```{r distribution-min-dist-ma}
ma_hist <- dn_lq_ma_mindist %>%
  ggplot(aes(x = closest)) +
  geom_histogram(binwidth = 10) +
  labs(title = "Shortest Distances Between Denny's and La Quinta Hotels",
       subtitle = "in Massachusetts",
       x = "Distance (km)") +
  theme_minimal()

dn_lq_ma_mindist %>%
  skim(closest)
```

### Exercise 11b

filter each data frame to just MA and then join. 
```{r ca-dn-lq}

dn_ca <- dennys %>%
  filter(state == "CA")

nrow(dn_ca)

lq_ca <- laquinta %>%
  filter(state == "CA")

nrow(lq_ca)

403*56 #prediction of rows

dn_lq_ca <- full_join(dn_ca, lq_ca, by = "state")
dn_lq_ca
```
```{r add-distance-ca}
dn_lq_ca <- dn_lq_ca %>%
   mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```

```{r min-dist-ca}

dn_lq_ca_mindist <- dn_lq_ca %>%
  group_by(address.x) %>%
  summarise(closest = min(distance)) %>%
  arrange(-desc(closest))
```

```{r distribution-min-dist-ca}
ca_hist <- dn_lq_ma_mindist %>%
  ggplot(aes(x = closest)) +
  geom_histogram(binwidth = 10) +
  labs(title = "Shortest Distances Between Denny's and La Quinta Hotels",
       subtitle = "in California",
       x = "Distance (km)") +
  theme_minimal()

dn_lq_ma_mindist %>%
  skim(closest)
```

```{r}
dn_lq_ma_mindist <- dn_lq_ma_mindist %>%
  mutate(state = "MA")
nrow(dn_lq_ma_mindist) 

dn_lq_nc_mindist <- dn_lq_nc_mindist %>%
  mutate(state = "NC")
nrow(dn_lq_nc_mindist)

dn_lq_tx_mindist <- dn_lq_tx_mindist %>%
  mutate(state = "TX")

dn_lq_ca_mindist <- dn_lq_ca_mindist %>%
mutate(state = "CA")


dn_lq_ma_nc_mindist <- full_join(dn_lq_ma_mindist, dn_lq_nc_mindist)

dn_lq_ma_nc_tx_mindist <- full_join(dn_lq_ma_nc_mindist,dn_lq_tx_mindist)

dn_lq_ma_nc_tx_ca_mindist <- full_join(dn_lq_ma_nc_tx_mindist, dn_lq_ca_mindist)

dn_lq_ma_nc_tx_ca_mindist_group <- dn_lq_ma_nc_tx_ca_mindist %>%
  group_by(state) %>%
  summarise(med = median(closest))

dn_lq_ma_nc_tx_ca_mindist %>%
  ggplot(aes(x = closest, color = state, fill = state)) +
  geom_density(alpha = 0.5) +
  geom_vline(data = dn_lq_ma_nc_tx_ca_mindist_group, aes(xintercept = med, color = state)) +
  labs(
    title = "Are Denny's and La Quinta Hotels close to one another?",
    subtitle = "in California, Massachusetts, North Carolina, and Texas",
    x = "Closest distance between a Denny's and a La Quinta Hotel (km)",
    y = "",
    caption = "Source: bit.ly/3bFr9o0 & bit.ly/2Oi5OIB"
  ) +
  scale_color_manual(values = c(
    "MA" = "#c60c30",
    "TX" = "#BF5700",
    "NC" = "#4B9CD3",
    "CA" = "#552583"
  )) +
  scale_fill_manual(values = c(
    "MA" = "#c60c30",
    "TX" = "#BF5700",
    "NC" = "#4B9CD3",
    "CA" = "#552583"
  )) +
  theme_minimal() +
  facet_wrap(~state, nrow = 4, ncol = 1)
  
```

```{r summary-stats-tx-nc-ma}

dn_lq_ma_nc_tx_ca_mindist %>%
  group_by(state) %>%
  skim(closest)

```

Mitch Hedberg's joke, "La Quinta is Spanish for "next to Denny's", most likely holds true in Texas compared to North Carolina and Massachusetts. 75 percent of La Quinta hotels are within 6.63 km of one another and on average are about 3.37km apart. 