---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "John Adams"
date: "3/12/21"
output: github_document
---

## Lab 4 Directions
https://rstudio-education.github.io/datascience-box/course-materials/lab-instructions/lab-04/lab-04-viz-sp-data.html

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
library(maps)
```

```{r load-data, message = FALSE}
states <- read_csv(url("https://raw.githubusercontent.com/rstudio-education/datascience-box/master/course-materials/starters/lab/lab-04-viz-sp-data/data/states.csv"))
```

### Exercise 1

```{r dennys-data-dimensions}
glimpse(dennys)
```

There are 1643 Denny's restaurants and 6 variables. 

### Exercise 2

```{r laquinta-data-dimensions}
glimpse(laquinta)
```

There are 909 La Quinta hotels and 6 variables. 

### Exercise 3

There aren't any Denny's outside of the US, but La Quinta does have hotels outside of the US. 

### Exercise 4

Using the data you could:
- visualize the location using long and lat 
- count the state variable

### Exercise 5

```{r denny-outside-us}

dennys %>%
  filter(!(state %in% states$abbreviation))

```

### Exercise 6

```{r dennys-with-country}
dn <- dennys %>%
        mutate(country = "United States")
```

### Exercise 8

```{r laquinta-with-country}

laquinta %>%
  filter(!(state %in% states$abbreviation))

lq <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state %in% c("AG", "QR", "CH", "NL", "VE", "PU", "SL" )  ~ "Mexico",
    state == "FM" ~ "Honduras"
  )) %>%
  filter(country == "United States")
  
```

```{r n-locations_state}

dn %>%
  group_by(state) %>%
  count() %>%
  arrange(desc(n))

lq %>%
  group_by(state) %>%
  count() %>%
  arrange(desc(n))
```

These numbers aren't that surprising as California, Texas, and Florida are all big states that attract a lot of tourists. 

### Exercise 9 

```{r joining-dn-state-data}
dn %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))

```
### Exercise 10

```{r dn-locations-sq-mile}
dn %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(
    area_thousands = area/1000,
    locations_per_area_thousands = n/(area/1000)) %>%
  arrange(desc(locations_per_area_thousands))

lq %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(
    area_thousands = area/1000,
    locations_per_area_thousands = n/(area/1000)) %>%
  arrange(desc(locations_per_area_thousands))

```

```{r merge-data}

dn <- dn %>%
  mutate(establishment = "Denny's")
lq <- lq %>%
  mutate(establishment = "La Quinta")

dn_lq <- bind_rows(dn, lq)

```

```{r plot-locations}

dn_lq %>%
  ggplot(aes(x = longitude, y = latitude, color = establishment)) +
  borders("state") +
  geom_point(alpha = 0.25, size = 0.5) +
  coord_quickmap() +
  labs(
    title = "Are Denny's Located Near La Quinta Hotels in the United States?",
    caption = "Source: bit.ly/3bFr9o0 & bit.ly/2Oi5OIB",
    x = NULL, 
    y = NULL,
    color = ""
  ) +
  scale_color_manual(values = c(
    "Denny's" = "#ffde24",
    "La Quinta" = "#005033"
  )) +
  theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = NA),
    axis.text = element_blank(),
    legend.position = "right", 
    axis.ticks = element_blank()
  ) 
```

### Trying to figure out how to map just one state

```{r nc-location-comp}
nc <- map_data('state', "north carolina")

mid_range <- function(x) mean(range(x))

seats <- do.call(rbind, lapply(split(nc, nc$subregion), function(d) {
  data.frame(lat = mid_range(d$lat), long = mid_range(d$long), subregion = unique(d$subregion))
}))


nc_dn_lq <- dn_lq  %>%
  filter(state == "NC")

nc <- nc %>%
  rename(longitude = long,
         latitude = lat
  )

nc_dn_lq_map <- full_join(nc_dn_lq, nc)

nc_dn_lq_map_poly <- full_join(nc_dn_lq, nc) %>%
filter(!is.na(group))
  
nc_dn_lq_map %>%  
  filter(is.na(group)) %>%
  ggplot(aes(x = longitude, y = latitude)) +
  geom_polygon(aes(group = group), data = nc_dn_lq_map_poly, fill = NA, colour = "grey60") +
  coord_quickmap() +
  geom_point(alpha = 1, size = 1, aes(color = establishment)) +
  labs(
    title = "Are Denny's Located Near La Quinta Hotels in the United States?",
    caption = "Source: bit.ly/3bFr9o0 & bit.ly/2Oi5OIB",
    x = NULL, 
    y = NULL,
    color = ""
  ) +
  scale_color_manual(values = c(
    "Denny's" = "#ffde24",
    "La Quinta" = "#005033"
  )) +
  theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = NA),
    axis.text = element_blank(),
    legend.position = "right", 
    axis.ticks = element_blank()
  ) 


```

### After figuring out how to map one state

North Carolina 

```{r final-nc-state}
  dn_lq %>%
    filter(state == "NC") %>%
  ggplot(aes(x = longitude, y = latitude)) +
  geom_polygon(aes(group = group), data = nc, fill = NA, colour = "grey60") +
  coord_quickmap() +
  geom_point(alpha = 1, size = 1, aes(color = establishment)) +
  labs(
    title = "Are Denny's Located Near La Quinta Hotels in North Carolina?",
    caption = "Source: bit.ly/3bFr9o0 & bit.ly/2Oi5OIB",
    x = NULL, 
    y = NULL,
    color = ""
  ) +
  scale_color_manual(values = c(
    "Denny's" = "#ffde24",
    "La Quinta" = "#005033"
  )) +
  theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = NA),
    axis.text = element_blank(),
    legend.position = "right", 
    axis.ticks = element_blank()
  ) 
```

```{r final-tx-state}
tx <- map_data('state', "texas")
mid_range <- function(x) mean(range(x))
seats_tx <- do.call(rbind, lapply(split(tx, tx$subregion), function(d) {
  data.frame(lat = mid_range(d$lat), long = mid_range(d$long), subregion = unique(d$subregion))
})) 

tx <- tx %>%
    rename(longitude = long,
         latitude = lat
  )

dn_lq %>%
    filter(state == "TX") %>%
  ggplot(aes(x = longitude, y = latitude)) +
  geom_polygon(aes(group = group), data = tx, fill = NA, colour = "grey60") +
  coord_quickmap() +
  geom_point(alpha = 0.5, size = 1, aes(color = establishment)) +
  labs(
    title = "Are Denny's Located Near La Quinta Hotels in Texas?",
    caption = "Source: bit.ly/3bFr9o0 & bit.ly/2Oi5OIB",
    x = NULL, 
    y = NULL,
    color = ""
  ) +
  scale_color_manual(values = c(
    "Denny's" = "#ffde24",
    "La Quinta" = "#005033"
  )) +
  theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = NA),
    axis.text = element_blank(),
    legend.position = "right", 
    axis.ticks = element_blank()
  ) 
```



