---
title: "Lab 03 - Nobel Laureates"
author: "John Adams"
date: "3/12/21"
output: github_document
---

## Lab directions

https://rstudio-education.github.io/datascience-box/course-materials/lab-instructions/lab-03/lab-03-nobel-laureates.html

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv(url("https://raw.githubusercontent.com/rstudio-education/datascience-box/master/course-materials/starters/lab/lab-03-nobel-laureates/data/nobel.csv"))
```

## Exercises

### Exercise 1

```{r get-to-know-data}

glimpse(nobel)

```
There are 935 nobel laureates and 26 variables. 

### Exercise 2

```{r nobel-living-data}
nobel_living <- nobel %>%
                filter(
                  !is.na(country),
                  gender != "org",
                  is.na(died_date)
                )
glimpse(nobel_living)

```


### Exercise 3

```{r data}
nobel_living <- nobel_living %>%
    mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )

nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
        
```

```{r faceted-bars-category}

nobel_living_science %>%
  ggplot(aes(x = country_us)) +
  geom_bar() +
  coord_flip() +
  labs(title = "Was the Nobel laurete in the US when he/she won?",
       subtitle = "For scientific Nobel prizes only",
       x = "",
       y = "") +
  facet_wrap(~category, nrow = 2)

```

### Exercise 4

```{r born-usa-data}

nobel_living_science <- nobel_living_science %>%
  mutate(
    born_country_us = if_else(born_country == "USA", "USA", "Other")
  )

```


### Exercise 5

```{r segmented-bars-in-us-born-us}
nobel_living_science %>%
  ggplot(aes(x = country_us, fill = born_country_us)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(title = "Was the Nobel laurete in the US when he/she won?",
       subtitle = "For scientific Nobel prizes only",
       x = "",
       y = "",
       fill = "Where was the laurete born?") +
  theme_minimal() +
 scale_fill_manual(values = c(
    "Other" = "#756bb1",
    "USA" = "gray"
  ))  +
  facet_wrap(~category, nrow = 2)
```

```{r tables}

nobel_living_science %>%
  group_by(category, country_us) %>%
  count(born_country_us) %>% 
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>%
  pivot_wider(names_from = born_country_us, values_from = proportion) 

```

### Exercise 6

The visualization above supports the Buzzfeed claim. Between 20 and 37.5 percent of nobel winners located in the US when they won were born in another country.  
What also becomes clear through this visualizations is that US born winners are not winning while in another country. 